Well-quasi-orderings on word languages

### 2024-08-05 src/introduction.tex
update intro

\section{Introduction}
\label{introduction:sec}

A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
$\preceq$ such that every infinite sequence $x_0 \preceq x_1 \preceq x_2
\preceq \ldots$ contains a pair $x_i \preceq x_j$ with $i < j$.
Well-quasi-orderings serve as a core combinatorial tool powering many
termination arguments, and in particular applies to the verification of
infinite state transition systems \cite{ABDU96,ABDU98}.

@@

\subparagraph{Contribution} In this paper, we characterize for which languages
$L \subseteq \Sigma^*$ the other natural orderins on word are
well-quasi-ordered. When languages are decided by simple enough machines (e.g.,
finite automata or pushdown automata), we will also provide \emph{decision
procedures} to check whether the language is well-quasi-ordered by the
considered relation.

>>>>

% LTeX: language=en-GB 
\section{Introduction}
\label{introduction:sec}

A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
$\preceq$ such that every infinite sequence $x_0, x_1, x_2 , \ldots$ contains a
pair $x_i \preceq x_j$ with $i < j$. Well-quasi-orderings serve as a core
combinatorial tool powering many termination arguments, and in particular
applies to the verification of infinite state transition systems
\cite{ABDU96,ABDU98}. 
In this paper, we will use an equivalent definition of
well-quasi-orderings based on the non-existence of infinite \intro{antichains}
-- sets of pairwise incomparable elements -- as well as the non-existence of
infinite \intro{decreasing sequences} -- sequences of elements $x_0, x_1, x_2,
\ldots$ such that $x_{i+1} \prec x_{i}$ --.

@@

\subparagraph{Contribution} In this paper, we characterize for which languages
$L \subseteq \Sigma^*$ the other natural orderings on word are
well-quasi-ordered. When languages are decided by simple enough machines (e.g.,
finite automata or pushdown automata), we will also provide \emph{decision
procedures} to check whether the language is well-quasi-ordered by the
considered relation. Quite surprisingly, for the \kl{prefix relation} and the
\kl{suffix relation}, the only languages that are \kl{well-quasi-ordered} are
the \emph{finite} unions of \kl{chains}, that is, the simplest possible
well-quasi-ordered sets.
